### 2017-03-18 dissipative_hamiltonian/chapters/5.conc.tex

\section{Conclusion} \label{sec:5}

Here comes the conclusion