### 2017-03-19 dissipative_hamiltonian/chapters/5.conc.tex

\section{Conclusion} \label{sec:5}

In this paper we present the Reduced Dissipative Hamiltonian method. The method preserves the symplectic structure of dissipative Hamiltonian systems and guarantees the correct dissipation of energy through time integration. The canonical heat bath, coupled with the full system, constructs a closed system of evolution, therefore a symplectic time integrator can be applied to preserve the system energy and stability over-long time integration. Moreover, the method allows the usage of a generic ortho-symplectic basis for model reduction.

The numerical simulations illustrate that the RDH method can preserve the system energy with significant more accuracy than other methods. Further they show that the hidden strings assure that the dissipation of energy in the reduce system mimics the dissipation of energy in the full system. This ensures that the local error do not accumulate over long-time integration.

All in all, the Reduced Dissipative Hamiltonian method is a promising model reduction technique for dissipative Hamiltonian systems and can save significant computational cost while preserving the symplectic structure and the dissipation structure of the full system.