### 2017-07-11 greedy_hamiltonian_second_revision/chapters/6.conclusion.tex

\section{Conclusion} \label{chap:Con:1}

In this paper, we present a greedy approach for the construction of a reduced system that preserves the geometric structure of Hamiltonian systems. An iteration of the greedy method comprises searching the parameter space using the error in the Hamiltonian, to find the best basis vectors that increase the overall accuracy of the reduced basis. We argue that for a compact subset with exponentially small Kolmogorov $n$-width we recover exponentially fast convergence of the greedy algorithm. For fast approximation of nonlinear terms, the basis obtained by the greedy was combined with a symplectic DEIM to construct {\edit a reduced system with a Hamiltonian that is arbitrary close to the Hamiltonian of the original system.}

%Since the reduced system obtained by the greedy method is Hamiltonian, it preserves stability, and was illustrated in the numerical results.

The numerical results demonstrate that the greedy method can save substantial computational cost in the offline stage as compared to alternative SVD-based techniques. Also since the reduced system obtained by the greedy method is Hamiltonian, the greedy method yields a stable reduced system. Symplectic DEIM effectively reduces computational cost of approximating nonlinear terms while preserving stability and symplectic structure. Hence, the greedy method is an efficient model reduction technique that provides an accurate and stable reduced system for large-scale parametric Hamiltonian systems.

 % it preserves stability, and was illustrated in the numerical results.
%We showed that the Hamiltonian is preserved in the 
%As the accuracy of the reduced system obtained by the greedy is comparable with the SVD-based methods, the greedy method is a promising dimension reduction technique for parametric Hamiltonian systems.